News Across Languages - Cross-Lingual Document Similarity and Event Tracking

### 2015-03-24 rupnik2015.tex
images

The term event is vague and ambiguous, but for the practical purposes we define it as ``any significant happening that is being reported about in the media''. Examples of events would include shooting down of the Malaysia Airlines plane over Ukraine on July 18th, 2014 and HSBC's admittance of aiding their clients in tax evasion on February 9th, 2015 (Figure \ref{}). Events such as these are covered by many articles and the question is how to find all the articles in different languages that are describing a single event.

As input we consider a stream of articles in different languages and a list of events. Our goal is to assign articles to their corresponding events. A priori we do not know the coverage of the articles, that is not all the events may be covered and we do not know that all the articles necessarily fit into one of the events.  To achieve our goal, we divide the task into two parts: clustering the articles according to topic (regardless of language) and matching the clusters to events.

@@

In Section~\ref{sec:models} we have described three main algorithms for identifying similar articles in different languages. These algorithms were $k$-means, LSI and CCA. To test which of these algorithms performs best we performed the following test. Using each algorithm, we analyzed all articles in Event Registry to find the most similar articles in other languages. The values of the cross-lingual article linking features (\texttt{linkCount} and \texttt{avgSimScore}) were then recomputed for all annotated learning examples. A dataset with only these two features and the label was constructed and used to test how accurate prediction models can be induced using the data. The results of the experiment are shown in Table~\ref{table:linkingEvalAlgos}. The table shows for each of the algorithms the obtained classification accuracy, precision and recall. We can see that the most informative features for cluster linking are obtained using the CCA algorithm, while $k$-means performs the worse. Firstly we verified the AMRR (average of mean reciprocal ranks over all language pairs) performance of approaches by holding out $15000$ aligned test documents and using $300000$ aligned documents
 as the training set. Figure~\ref{pic:AMRR} shows AMRR score as the function of the number of feature vectors. It is clear that CCA outperforms LSI approach and K-means lags far behind.
Furthermore we inspected how number of features influences the accuracy of cluster linking. As we can see from Table~\ref{table:linkingEvalAlgos} choosing number of features vectors larger than $500$ barely effects linking performance. That is probably due to overfitting. Moreover we tested the performance of detecting large/small cluster linking, see Table~\ref{table:linkingEvalAlgosLargeSmall}.  As expected we obtained that it is harder to link small clusters and we need better retrieval accuracy in such cases. CCA outperforms other methods in this sense also, probably as it has the best retrieval performance, see Figure~\ref{pic:AMRR}. This can be also used
to get representatives of clusters more efficiently.

@@

\subsection{Scaling}
\emph{I think that this section should be short.}
100 languages!
%Memory 100 GB per day
%Newsfeed volume 430k articles per day, 210k english, 36k Spanish, 27k German, 17k French, ....
%Number of similarities, 430k similarities every 200ms, currently could handle 1M articles,
%(n \over 2) 500X500 proxy matrices, 300000x500 Projector matrices, 430000x500 data
%Premultiply with proxy matrices, compare blockwise!
One of the main advantages of our approach is that scales really well. It is fast, very robust to quality of training data, easily extendable, simple to implement and with low machine requirements. The similarity pipeline is the most demanding part and currently runs on two Intel Xeon E5-2667 v2, 3.30GHz processor machine with 256GB of RAM. That is enough to do similarity computation over 150 languages if needed. It currently uses Wikipedia as freely available knowledge base.
Similarity pipeline dramastically reduces the search space when linking clusters.

Currently we compute similarities over $24$ languages with tags: eng, spa, deu, zho, ita, fra, rus, swe, nld, tur, jpn, por, ara, fin, ron, kor, hrv, tam, hun, slv, pol, srp, cat, ukr but we support up to $100$ top Wikipedis languages. Our data stream is Newsfeed (http://newsfeed.ijs.si/) that produces $430k$ unique articles per day. Our system currently computes 400k similiarites per 200ms, that means that we compute $16*10^10$ similarities per day. We
store one day buffer for every language, for that we need 1.5GB of memory, we store documents as 500 dimensional vectors. We must note that time complexity of similarity computations scales linearly with dimension of the feature space and does not  depend of number of languages. For each article we compute top $10$  most similar ones in every other language.
For all linear algebra matrix and vector operations we use high performance numerical libraries as BLAS, OPENBLAS and Intel MKL. This currently allows us to process more than million articles per day.
In our current implementation we use variation of hub approach. Our projector matrices are of size $500\times 300000,$ so every projector takes about $1.1$ GB of RAM. Moreover we need proxy matrices of size $500x500$ for every language pair. That is 0.5 GB for $24$ languages and $9.2$ GB for $100$ languages. All together we need around 135 GB of RAM for the system with 100 languages. Using proxy matrices we can project all the input documents in the common space. That enables us to do blockwise similarity computations and
additional improvement in our system efficiency. Our code can be therefore easily parallelized using matrix multiplication instead of doing more matrix - vector multiplications. This speeds up our code by factor around $4.$ In this way we obtain some caching gains and ability to use vectorization.
Our system is also easily extendable. If we want to add new language we only need to compute a projector matrix and proxy matrices with all other already available languages.

Add the conclusion of the story and the wrap up, Andrej needs help with this.

>>>>

The term event is vague and ambiguous, but for the practical purposes we define it as ``any significant happening that is being reported about in the media''. Examples of events would include shooting down of the Malaysia Airlines plane over Ukraine on July 18th, 2014 and HSBC's admittance of aiding their clients in tax evasion on February 9th, 2015 (Figure \ref{fig:event2}). Events such as these are covered by many articles and the question is how to find all the articles in different languages that are describing a single event.

\begin{figure}
\end{figure}
As input we consider a stream of articles in different languages and a list of events. Our goal is to assign articles to their corresponding events. A priori we do not know the coverage of the articles, that is not all the events may be covered and we do not know that all the articles necessarily fit into one of the events.  To achieve our goal, we divide the task into two parts: clustering the articles according to topic (regardless of language) and matching the clusters to events.

@@

In Section~\ref{sec:models} we have described three main algorithms for identifying similar articles in different languages. These algorithms were $k$-means, LSI and CCA. To test which of these algorithms performs best we performed the following test. Using each algorithm, we analyzed all articles in Event Registry to find the most similar articles in other languages. The values of the cross-lingual article linking features (\texttt{linkCount} and \texttt{avgSimScore}) were then recomputed for all annotated learning examples. A dataset with only these two features and the label was constructed and used to test how accurate prediction models can be induced using the data. The results of the experiment are shown in Table~\ref{table:linkingEvalAlgos}. The table shows for each of the algorithms the obtained classification accuracy, precision and recall. We can see that the most informative features for cluster linking are obtained using the CCA algorithm, while $k$-means performs the worse. Firstly we verified the AMRR (average of mean reciprocal ranks over all language pairs) performance of approaches by holding out $15000$ aligned test documents and using $300000$ aligned documents
 as the training set. Figure~\ref{pc:AMRR} shows AMRR score as the function of the number of feature vectors. It is clear that CCA outperforms LSI approach and K-means lags far behind.
Furthermore we inspected how number of features influences the accuracy of cluster linking. As we can see from Table~\ref{table:linkingEvalAlgos} choosing number of features vectors larger than $500$ barely effects linking performance. That is probably due to overfitting. Moreover we tested the performance of detecting large/small cluster linking, see Table~\ref{table:linkingEvalAlgosLargeSmall}.  As expected we obtained that it is harder to link small clusters and we need better retrieval. Hub CCA outperforms other methods in this sense also, probably as it has the best retrieval performance, see Figure~\ref{pic:AMRR}.

@@

\subsection{Scaling}
\emph{I think that this section should be short.}
100 languages!
%Memory 100 GB per day
%Newsfeed volume 430k articles per day, 210k english, 36k Spanish, 27k German, 17k French, ....
%Number of similarities, 430k similarities every 200ms, currently could handle 1M articles,
%(n \over 2) 500X500 proxy matrices, 300000x500 Projector matrices, 430000x500 data
%Premultiply with proxy matrices, compare blockwise!
One of the main advantages of our approach is that scales really well. It is fast, very robust to quality of training data, easily extendable, simple to implement and with low machine requirements. Ut currently runs on two Intel Xeon E5-2667 v2, 3.30GHz processor machine with 256GB of RAM. That is enough to do similarity computation over 150 languages if needed. It currently uses Wikipedia as freely available knowledge base.

Currently we compute similarities over $24$ languages with tags: eng, spa, deu, zho, ita, fra, rus, swe, nld, tur, jpn, por, ara, fin, ron, kor, hrv, tam, hun, slv, pol, srp, cat, ukr but we support up to $100$ top Wikipedis languages. Our data stream is Newsfeed(http://newsfeed.ijs.si/) that produces $430k$ unique articles per day. Our system currently computes 400k similiarites per 200ms, that means that we compute $16*10^10$ similarities per day. We
store one day buffer for every language, for that we need 1.5GB of memory, we store documents as 500 dimensional vectors. For each article we compute top $10$ most similar ones in every language.
For all linear algebra matrix and vector operations we use high performance numerical libraries as BLAS, OPENBLAS and Intel MKL. This currently allows us to process more than million articles per day.
In our current implementation we use variation of hub approach. Our projector matrices are of size $500\times 300000,$ so every projector takes about $1.1$ GB of RAM. Moreover we need proxy matrices of size $500x500$ for every language pair. That is 0.5 GB for $24$ languages and $9.2$ GB for $100$ languages. All together we need around 135 GB of RAM for the system with 100 languages. Using proxy matrices we can project all the input documents in the common space. That enables us to do blockwise similarity computations and
additional improvement in our system efficiency. Our code can be therefore easily parallelized using matrix multiplication instead of doing more matrix - vector multiplications. This speeds up our code by factor around $4.$ In this way we obtain some caching gains and ability to use vectorization.
Our system is also easily extendable. If we want to add a language we only need to compute a projector matrix and proxy matrices with all other already available languages.

Comment on the caching gains in our implementation.¸¸Add the conclusion if the story.
